\Large \bf Uncovering Distributed Storage System Bugs during\\Testing (not Production!)

### 2015-09-21 experience.tex
Edits to MigratingTable case study description.

The Live Azure Table Migration (MigratingTable) is a library for \emph{transparently migrating} a data set between tables in an Azure Storage service \emph{while} an application is accessing this data set. This case study is particularly interesting because its developers \emph{co-developed} the system with its \psharp model to increase confidence in the former's implementation. Indeed, this co-development process resulted in many tricky bugs discovered before the system went into production (see \S\ref{sec:eval}). The MigratingTable modeling effort differs from the vNext case study: the vNext developers focused on specifying a single liveness property; whereas the MigratingTable developers had to write many complex safety properties. MigratingTable is also interesting because it heavily uses the \texttt{async} and \texttt{await} \csharp 5.0 primitives, which were not originally supported in \psharp. To be able to test MigratingTable, as well as other systems written in \csharp 5.0, we had to extend \psharp with support for intra-machine concurrency (see \S\ref{sec:psharp:testing}).

@@

MigratingTable provides a \emph{virtual table} that has a similar interface to an ordinary Azure table. This interface is named \texttt{IChainTable}. The virtual table is backed by a pair of \emph{old} and \emph{new} tables. A background \emph{migrator} job is responsible for moving all data from the old table to the new table. Meanwhile, each read and write operation issued to the virtual table is translated to a sequence of reads and writes on the backend tables according to a protocol that guarantees linearizability of operations on the virtual table across multiple application processes, assuming that the backend tables respect their own linearizability guarantees.

The testing goal was to ensure that when multiple application processes, implementing the \texttt{IChainTable} interface, issue read and write operations to their own MigratingTable instances with the same backend tables, the behavior complies with the specification of \texttt{IChainTable} for the combined \emph{input history}.

@@

The main challenge behind testing MigratingTable is that there are many possible input histories and that the system is highly concurrent. The developers could have tested specific input histories, but they were not confident that this approach would be effective in catching bugs, especially because concurrency increases the potential for difficult-to-foresee interactions between seemingly independent parts of the code.

\textbf{Modeling and testing}.
Towards testing the MigratingTable library, the developers wrote an in-memory reference implementation of the \texttt{IChainTable} interface, called \texttt{SpecTable}, which can be used for comparing the output of MigratingTable on an arbitrary input history. This enabled sampling from a distribution that was defined over all possible input histories within certain bounds. The \psharp systematic testing engine automatically takes care of enumerating different input histories. The MigratingTable was instrumented to report the intended \emph{linearization point} of each input call. Specifically, after each input call completes, MigratingTable reports whether that call was the linearization point, which may depend on the result of the call.  This makes it possible to check the correctness property as the model executes.

@@

The \psharp test harness of MigratingTable consists of a \texttt{Tables} machine containing the old, new and reference table implementations; a collection of \texttt{Service} machines containing identically configured MigratingTables; and a \texttt{Migrator} machine that performs the background migration (see Figure~\ref{fig:mockedmigration}). Each \texttt{Service} machine issues a random sequence of input calls to its MigratingTable, which sends backend calls to the \texttt{Tables} machine. When MigratingTable reports the linearization point of an input call, the \texttt{Service} machine sends that input call to the reference table.  When an input call completes, the \texttt{Service} machine checks that the results from the MigratingTable and the reference table agree.

\psharp captures and controls the interleaving of the backend calls. To ensure that the reference table is never observed to be out of sync with the backend tables, after the \texttt{Tables} machine processes a backend call, it enters a state that defers further backend calls until MigratingTable has reported whether the backend call was a linearization point and (if so) the call to the reference table has been made.

>>>>

The Live Azure Table Migration library (MigratingTable) is designed to transparently migrate a data set between tables in an Azure Storage service while an application is accessing this data set. This case study is particularly interesting because its developers \emph{co-developed} the system with its \psharp test to increase confidence in the system. Indeed, this co-development process resulted in the discovery of several tricky bugs before the system went into production (see \S\ref{sec:eval}). The MigratingTable testing effort differs from the vNext case study: the vNext developers focused on specifying a single liveness property, while the MigratingTable test checks complete compliance with an interface specification. MigratingTable is also interesting because it heavily uses the \texttt{async} and \texttt{await} \csharp 5.0 primitives, which were not originally supported in \psharp. To be able to test MigratingTable, as well as other systems written in \csharp 5.0, we had to extend \psharp with support for intra-machine concurrency (see \S\ref{sec:psharp:testing}).

@@

MigratingTable provides a \emph{virtual table} that has a similar interface to an ordinary Azure table. This interface is named \texttt{IChainTable}. The virtual table is backed by a pair of \emph{old} and \emph{new} tables, which must also implement \texttt{IChainTable}. A background \emph{migrator} job is responsible for moving all data from the old table to the new table. Meanwhile, each read and write operation issued to the virtual table is translated to a sequence of reads and writes on the backend tables according to a protocol, which is intended to guarantee that when multiple application processes issue \emph{input} operations to their own MigratingTable instances with the same backend tables, the output of the MigratingTable instances is compliant with the specification of \texttt{IChainTable} for the combined \emph{input history} as if all the operations were done on a single table. The goal of the \psharp test was to verify this property.

@@

\textbf{Modeling and testing}.
The main challenge behind testing MigratingTable is that there are many possible input histories and that the system is highly concurrent. The developers could have chosen specific input histories to test, but they were not confident that this approach would be effective in catching bugs, especially because concurrency increases the potential for difficult-to-foresee interactions between seemingly independent parts of the code. Instead they took advantage of the simplicity and near-determinism of the \texttt{IChainTable} specification to write an in-memory reference implementation, called \texttt{SpecTable}, to which the output of MigratingTable on an arbitrary input history can be compared. Then the test could sample from a distribution that was defined over all possible input histories within certain bounds, so that under the \emph{small scope hypothesis} that any given bug in MigratingTable leads to incorrect output for at least one input history in the distribution, each iteration of the test would have a positive probability of detecting the bug.

The MigratingTable was instrumented to report the intended \emph{linearization point} of each input operation on the abstract, shared virtual table. This allows the test harness to keep a \emph{reference} \texttt{SpecTable} in sync with the virtual table by sending each operation to the reference table when the linearization point on the virtual table is reached. The output of the operation on the MigratingTable is then checked against its output on the reference table.

@@

The \psharp test harness of MigratingTable consists of a \texttt{Tables} machine containing the old and new backend tables and the reference table; a collection of \texttt{Service} machines containing identically configured MigratingTables; and a \texttt{Migrator} machine that performs the background migration (see Figure~\ref{fig:mockedmigration}). Each \texttt{Service} machine issues a random sequence of input operations to its MigratingTable, which sends backend operations to the \texttt{Tables} machine. The \texttt{Service} machine sends the operation to the reference table at the linearization point and compares the results as previously described.

% This paragraph can probably be removed if we are out of space. ~ Matt 2015-09-20
\psharp captures and controls the interleaving of the backend operations. To ensure that the reference table is never observed to be out of sync with the backend tables, after the \texttt{Tables} machine processes a backend operation, it enters a state that defers further backend operations until MigratingTable has reported whether the backend operation was a linearization point and (if so) the operation on the reference table has been completed.
